---
title: "10XMultiome"
output: html_document
date: "2024-01-04"
---

Loading libraries
```{r}
library(Signac)
library(Seurat)
library(SeuratDisk)
library(SeuratObject)
library(SeuratData)
library(GenomicRanges)
library(future)
library(EnsDb.Hsapiens.v86)
library(Azimuth)
library(BSgenome.Hsapiens.UCSC.hg38)

# setRepositories(ind = 1:3, addURLs = c('https://satijalab.r-universe.dev', 'https://bnprks.r-universe.dev/'))
# install.packages(c("BPCells", "presto", "glmGamPoi"))
# BiocManager::install("TFBSTools", type = "source", force = TRUE)

# plan("multisession", workers = 64)
# options(future.globals.maxSize = 60000 * 1024^2) # for 60 Gb RAM
```

Merging objects
```{r}
# Load the ATAC and RNA objects
pbmc.atac <- readRDS("/home/tjq4954/Multiome_PBMC/Data/atac/pbmc.atac.rds")
pbmc.rna <- readRDS("/home/tjq4954/Multiome_PBMC/Data/rna/pbmc.rna.rds")

# Check if the cell names are formatted the same
identical(colnames(pbmc.atac), colnames(pbmc.rna))

# Merging objects
pbmc.rna[['ATAC']] <- pbmc.atac[['ATAC']] # assay
pbmc.rna[['lsi']] <- pbmc.atac[['lsi']] # embedding
pbmc.rna@meta.data <- cbind(pbmc.rna@meta.data, pbmc.atac@meta.data)
pbmc <- pbmc.rna

# Specify the full path with the filename for the RDS file
save_path <- "/home/tjq4954/Multiome_PBMC/Data/pbmc.rds"
# Save the Seurat object as an RDS file
saveRDS(pbmc, file = save_path)
```

Quality control
```{r}
library(gridExtra)
setwd("/home/tjq4954/Multiome_PBMC/Data")
pbmc <- readRDS("pbmc.rds")

DefaultAssay(pbmc) <- "ATAC"

# Create some important QC metrics
pbmc <- NucleosomeSignal(pbmc)
pbmc <- TSSEnrichment(pbmc)

VlnPlot(
  object = pbmc,
  features = c("nCount_RNA", "nCount_ATAC","TSS.enrichment","nucleosome_signal","percent.mt"),
  ncol = 3,
  pt.size = 0
)

# filter out low quality cells
pbmc <- subset(
  x = pbmc,
  subset = nCount_ATAC < 100000 &
    nCount_RNA < 25000 &
    nCount_ATAC > 1000 &
    nCount_RNA > 1000 &
    nucleosome_signal < 2 &
    TSS.enrichment > 1
)

# Create the contingency table
Idents(pbmc) <- "predicted.celltype.l2"
celltype_ident_table <- table(pbmc$predicted.celltype.l2, pbmc$orig.ident)
table_df <- as.data.frame.matrix(celltype_ident_table)
write.csv(table_df, "/home/tjq4954/Multiome_PBMC/Visuals/celltype_ident_table.csv", row.names = TRUE)


# Specify the full path with the filename for the RDS file
save_path <- "/home/tjq4954/Multiome_PBMC/Data/pbmc_subset.rds"
# Save the Seurat object as an RDS file
saveRDS(pbmc, file = save_path)
```

Peak calling
```{r}
# devtools::install_github("GreenleafLab/ArchR", ref="master", repos = BiocManager::repositories())
# library(ArchR)
# ArchR::installExtraPackages()

setwd("/home/tjq4954/Multiome_PBMC/Data")
pbmc <- readRDS("pbmc_subset.rds")

# Assuming you have already identified which samples are cases and controls
control_samples <- c("RNA_280", "RNA_302")
case_samples <- c("RNA_72", "RNA_203", "RNA_271", "RNA_294")

# Subset pbmc object into case and control groups
pbmc_control <- subset(pbmc, subset = orig.ident %in% control_samples)
pbmc_case <- subset(pbmc, subset = orig.ident %in% case_samples)

# call peaks using MACS
peaks_control <- CallPeaks(pbmc_control
                           # ,group.by = "predicted.celltype.l2"
                           )
peaks_case <- CallPeaks(pbmc_case
                        # ,group.by = "predicted.celltype.l2"
                        )

# remove peaks on nonstandard chromosomes and in genomic blacklist regions
peaks_control <- keepStandardChromosomes(peaks_control, pruning.mode = "coarse")
peaks_control <- subsetByOverlaps(x = peaks_control, ranges = blacklist_hg38_unified, invert = TRUE)
peaks_case <- keepStandardChromosomes(peaks_case, pruning.mode = "coarse")
peaks_case <- subsetByOverlaps(x = peaks_case, ranges = blacklist_hg38_unified, invert = TRUE)

# quantify counts in each peak
macs2_counts_control <- FeatureMatrix(
  fragments = Fragments(pbmc_control),
  features = peaks_control,
  cells = colnames(pbmc_control)
)
macs2_counts_case <- FeatureMatrix(
  fragments = Fragments(pbmc_case),
  features = peaks_case,
  cells = colnames(pbmc_case)
)

# create a new assay using the MACS2 peak set and add it to the Seurat object
pbmc_control[["peaks"]] <- CreateChromatinAssay(
  counts = macs2_counts_control,
  fragments = Fragments(pbmc_control),
  annotation = annotations
)
pbmc_case[["peaks"]] <- CreateChromatinAssay(
  counts = macs2_counts_case,
  fragments = Fragments(pbmc_case),
  annotation = annotations
)

# Specify the full path with the filename for the RDS file
save_path_case <- "/home/tjq4954/Multiome_PBMC/Data/pbmc_peaks_case.rds"
# Save the Seurat object as an RDS file
saveRDS(pbmc_case, file = save_path_case)

# Specify the full path with the filename for the RDS file
save_path_control <- "/home/tjq4954/Multiome_PBMC/Data/pbmc_peaks_control.rds"
# Save the Seurat object as an RDS file
saveRDS(pbmc_control, file = save_path_control)

# # Specify the full path with the filename for the RDS file
# save_path <- "/home/tjq4954/Multiome_PBMC/Data/pbmc_peaks.rds"
# # Save the Seurat object as an RDS file
# saveRDS(pbmc, file = save_path)
```

DEG & DA Analysis
```{r}
# setwd("/home/tjq4954/Multiome_PBMC/Data")
# pbmc <- readRDS("pbmc_peaks.rds")

# setwd("/home/tjq4954/Multiome_PBMC/Data")
# pbmc_control <- readRDS("pbmc_peaks_control.rds")
# setwd("/home/tjq4954/Multiome_PBMC/Data")
# pbmc_case <- readRDS("pbmc_peaks_case.rds")

# DefaultAssay(pbmc) <- "RNA" # already performed in the RNA integration script
# pbmc <- SCTransform(pbmc)
# pbmc <- RunPCA(pbmc)

DefaultAssay(pbmc_case) <- "peaks"
pbmc_case <- FindTopFeatures(pbmc_case, min.cutoff = 5)
pbmc_case <- RunTFIDF(pbmc_case)
pbmc_case <- RunSVD(pbmc_case)

DefaultAssay(pbmc_control) <- "peaks"
pbmc_control <- FindTopFeatures(pbmc_control, min.cutoff = 5)
pbmc_control <- RunTFIDF(pbmc_control)
pbmc_control <- RunSVD(pbmc_control)

```

Joint UMAP Visualization
```{r}
library(ggplot2)

# set a reasonable order for cell types to be displayed when plotting
# levels(pbmc) <- c("CD4 Naive", "CD4 TCM", "CD4 CTL", "CD4 TEM", "CD4 Proliferating",
#                   "CD8 Naive", "dnT",
#                  "CD8 TEM", "CD8 TCM", "CD8 Proliferating", "MAIT", "NK", "NK_CD56bright",
#                  "NK Proliferating", "gdT",
#                  "Treg", "B naive", "B intermediate", "B memory", "Plasmablast",
#                  "CD14 Mono", "CD16 Mono",
#                  "cDC1", "cDC2", "pDC", "HSPC", "Eryth", "ASDC", "ILC", "Platelet")

# Remove platelets and erythrocytes
pbmc <- subset(pbmc, idents = c("Eryth", "Platelet"), invert = TRUE)
unique(pbmc$predicted.celltype.l2)

# build a joint neighbor graph using both assays
pbmc <- FindMultiModalNeighbors(
  object = pbmc,
  reduction.list = list("pca", "lsi"), 
  dims.list = list(1:50, 2:40),
  modality.weight.name = "RNA.weight",
  verbose = TRUE
)

# build a joint UMAP visualization
pbmc <- RunUMAP(
  object = pbmc,
  nn.name = "weighted.nn",
  assay = "RNA",
  verbose = TRUE
)

# WNN UMAP - Cell Type
plot1 <- DimPlot(pbmc, label = TRUE, repel = TRUE, reduction = "umap", group.by = "predicted.celltype.l2") + NoLegend() + labs(title = "Joint UMAP")
ggsave(filename = "/home/tjq4954/Multiome_PBMC/Visuals/JointUMAP_CellType.png", plot = plot1, width = 10, height = 8, dpi = 600)
# WNN UMAP - Sample ID
plot2 <- DimPlot(pbmc, label = TRUE, repel = TRUE, reduction = "umap", group.by = "orig.ident") + NoLegend()
ggsave(filename = "/home/tjq4954/Multiome_PBMC/Visuals/JointUMAP_SampleID.png", plot = plot2, width = 10, height = 8, dpi = 600)
```

Coverage Plot
```{r}
# Research Idea #1: PSMB8-AS1 induced PSMB9 transcription by recruiting the transcription factor NONO (non-POU domain-containing octamer-binding protein) and binding to the PSMB9 promoter. PSMB9 (proteasome subunit-β type-9) elevated VCAM1 and ICAM1 expression via the upregulation of ZEB1 (zinc finger E-box-binding homeobox 1).
# Research Idea #2: An interesting finding at first glance is the 4-5-log2 fold change elevation in cases (4 with CAD) versus controls (3 closely matched without CAD) in monocyte expression of c17orf99, which is IL-40, implicated in B cell activation and some inflammatory conditions.
# first compute the GC content for each peak
pbmc_control <- RegionStats(pbmc_control, genome = BSgenome.Hsapiens.UCSC.hg38)
pbmc_case <- RegionStats(pbmc_case, genome = BSgenome.Hsapiens.UCSC.hg38)

# link peaks to genes
pbmc_control <- LinkPeaks(
  object = pbmc_control,
  peak.assay = "peaks",
  expression.assay = "SCT",
  genes.use = c("C17orf99", "HIF1A", "IL1B", "IL10"),
  pvalue_cutoff = 0.2,
  score_cutoff = 0.2
)

pbmc_case <- LinkPeaks(
  object = pbmc_case,
  peak.assay = "peaks",
  expression.assay = "SCT",
  genes.use = c("C17orf99", "HIF1A", "IL1B", "IL10"),
  pvalue_cutoff = 0.2,
  score_cutoff = 0.2,
)

Links(pbmc_control)
Links(pbmc_case)

# Look up genes to plot
grep("C17orf99", rownames(pbmc@assays$SCT), value = TRUE)

p1 <- CoveragePlot(
  object = pbmc,
  region = "PSMB9",
  features = "PSMB9",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l2",
  idents = c("CD16 Mono","CD14 Mono")
)

p2 <- CoveragePlot(
  object = pbmc,
  region = "C17orf99",
  features = "C17orf99",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l2",
  idents = c("CD16 Mono","CD14 Mono"),
  extend.downstream = 5000
)

p3 <- CoveragePlot(
  object = pbmc,
  region = "IL1B",
  features = "IL1B",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l2",
  idents = c("CD16 Mono","CD14 Mono")
)

p4 <- CoveragePlot(
  object = pbmc,
  region = "LYZ",
  features = "LYZ",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l2",
  idents = c("CD16 Mono","CD14 Mono")
  # , extend.upstream = 5000,
  # extend.downstream = 5000
)

patchwork::wrap_plots(p1,p2, ncol = 1)
patchwork::wrap_plots(p3,p4, ncol = 1)

```

Cases vs. Controls for IL1B, CCL4, MIF, HIF1A
```{r}
# Look up genes to plot
grep("C17orf99", rownames(pbmc@assays$SCT), value = TRUE)
# "PSMB8-AS1", "PSMB8", "PSMB9", "APOE", "VCAM1", "ICAM1", "ZEB1", "IL1B", "IL10"

# Coverage plots for case group
p1_case <- CoveragePlot(
  object = pbmc_case,
  region = "PSMB9",
  features = "PSMB9",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l1",
  # idents = c("CD16 Mono","CD14 Mono"),
  extend.upstream = 5000,
  extend.downstream = 5000,
  links = FALSE
)

p2_case <- CoveragePlot(
  object = pbmc_case,
  region = "C17orf99",
  features = "C17orf99",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l1",
  # idents = c("CD16 Mono","CD14 Mono"),
  extend.downstream = 5000,
  links = FALSE
)

p3_case <- CoveragePlot(
  object = pbmc_case,
  region = "IL1B",
  features = "IL1B",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l1",
  idents = c("Mono","DC"),
  links = FALSE
)

p4_case <- CoveragePlot(
  object = pbmc_case,
  region = "IL10",
  features = "IL10",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l1",
  # idents = c("CD16 Mono","CD14 Mono"),
  extend.downstream = 8000,
  links = FALSE
)

p5_case <- CoveragePlot(
  object = pbmc_case,
  region = "HIF1A",
  features = "HIF1A",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l1",
  # idents = c("CD16 Mono","CD14 Mono"),
  extend.downstream = 8000,
  links = FALSE
)

p6_case <- CoveragePlot(
  object = pbmc_case,
  region = "LYZ",
  features = "LYZ",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l1",
  # idents = c("CD16 Mono","CD14 Mono"),
  extend.upstream = 5000,
  extend.downstream = 5000,
  links = FALSE
)

# Coverage plots for control group
p1_control <- CoveragePlot(
  object = pbmc_control,
  region = "PSMB9",
  features = "PSMB9",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l1",
  # idents = c("CD16 Mono","CD14 Mono"),
  extend.upstream = 5000,
  extend.downstream = 5000,
  links = FALSE
)

p2_control <- CoveragePlot(
  object = pbmc_control,
  region = "C17orf99",
  features = "C17orf99",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l1",
  # idents = c("CD16 Mono","CD14 Mono"),
  extend.downstream = 5000,
  links = FALSE
)

p3_control <- CoveragePlot(
  object = pbmc_control,
  region = "IL1B",
  features = "IL1B",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l1",
  idents = c("Mono","DC"),
  links = FALSE
)

p4_control <- CoveragePlot(
  object = pbmc_control,
  region = "IL10",
  features = "IL10",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l1",
  # idents = c("CD16 Mono","CD14 Mono"),
  extend.downstream = 8000,
  links = FALSE
)

p5_control <- CoveragePlot(
  object = pbmc_control,
  region = "HIF1A",
  features = "HIF1A",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l1",
  # idents = c("CD16 Mono","CD14 Mono"),
  extend.downstream = 8000,
  links = FALSE
)

p6_control <- CoveragePlot(
  object = pbmc_control,
  region = "LYZ",
  features = "LYZ",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l1",
  # idents = c("CD16 Mono","CD14 Mono"),
  extend.upstream = 5000,
  extend.downstream = 5000,
  links = FALSE
)

# Add titles to case and control coverage plots
library(patchwork)
p1_case <- p1_case + plot_annotation(title = "PSMB9 - Case (Top) v.s. Control (Bottom)")
p2_case <- p2_case + plot_annotation(title = "C17orf99 (IL40) - Case (Top) v.s. Control (Bottom)")
p3_case <- p3_case + plot_annotation(title = "IL1B - Case (Top) v.s. Control (Bottom)")
p4_case <- p4_case + plot_annotation(title = "IL10 - Case (Top) v.s. Control (Bottom)")
p5_case <- p5_case + plot_annotation(title = "HIF1A - Case (Top) v.s. Control (Bottom)")
p6_case <- p6_case + plot_annotation(title = "LYZ - Case (Top) v.s. Control (Bottom)")

# Combine case and control plots for each gene
combined_p1 <- p1_case + p1_control + plot_layout(ncol = 1)
combined_p2 <- p2_case + p2_control + plot_layout(ncol = 1)
combined_p3 <- p3_case + p3_control + plot_layout(ncol = 1)
combined_p4 <- p4_case + p4_control + plot_layout(ncol = 1)
combined_p5 <- p5_case + p5_control + plot_layout(ncol = 1)
combined_p6 <- p6_case + p6_control + plot_layout(ncol = 1)

# Save the combined plots
ggsave(filename = "/home/tjq4954/Multiome_PBMC/Visuals/CoveragePlot_PSBM9.png", plot = combined_p1, width = 9, height = 15, dpi = 600)
ggsave(filename = "/home/tjq4954/Multiome_PBMC/Visuals/CoveragePlot_IL40.png", plot = combined_p2, width = 9, height = 15, dpi = 600)
ggsave(filename = "/home/tjq4954/Multiome_PBMC/Visuals/CoveragePlot_IL1B.png", plot = combined_p3, width = 7, height = 8, dpi = 600)
ggsave(filename = "/home/tjq4954/Multiome_PBMC/Visuals/CoveragePlot_IL10.png", plot = combined_p4, width = 9, height = 15, dpi = 600)
ggsave(filename = "/home/tjq4954/Multiome_PBMC/Visuals/CoveragePlot_HIF1A.png", plot = combined_p5, width = 9, height = 15, dpi = 600)
ggsave(filename = "/home/tjq4954/Multiome_PBMC/Visuals/CoveragePlot_LYZ.png", plot = combined_p6, width = 9, height = 15, dpi = 600)

# table(pbmc$orig.ident)
```

Coverage Plot with Peak Links
```{r}
# Customized Coverage Plot for HIF1A (Case)
cov_plot_case <- CoveragePlot(object = pbmc_case, region = "chr14-61490000-61770000", annotation = F, peaks = F, links =F)
gene_plot_case <- AnnotationPlot(object = pbmc_case, region = "chr14-61490000-61770000")
peak_plot_case <- PeakPlot(object = pbmc_case, region = "chr14-61490000-61770000")
link_plot_case <- LinkPlot(object=pbmc_case, region = "chr14-61490000-61770000")
expr_plot_case <- ExpressionPlot(object = pbmc_case, features = "HIF1A", assay = "SCT")
HIF1A_case <- CombineTracks(
  plotlist = list(cov_plot_case, peak_plot_case, gene_plot_case, link_plot_case),
  expression.plot = expr_plot_case,
  heights = c(10, 1, 2, 3),
  widths = c(10, 1)
)
HIF1A_case <- HIF1A_case + plot_annotation(title = "HIF1A Coverage Plot with Peak Links - Case")

# Customized Coverage Plot for HIF1A (Control)
cov_plot_control <- CoveragePlot(object = pbmc_control, region = "chr14-61490000-61770000", annotation = F, peaks = F, links =F)
gene_plot_control <- AnnotationPlot(object = pbmc_control, region = "chr14-61490000-61770000")
peak_plot_control <- PeakPlot(object = pbmc_control, region = "chr14-61490000-61770000")
link_plot_control <- LinkPlot(object=pbmc_control, region = "chr14-61490000-61770000")
expr_plot_control <- ExpressionPlot(object = pbmc_control, features = "HIF1A", assay = "SCT")
HIF1A_control <- CombineTracks(
  plotlist = list(cov_plot_control, peak_plot_control, gene_plot_control, link_plot_control),
  expression.plot = expr_plot_control,
  heights = c(10, 1, 2, 3),
  widths = c(10, 1)
)
HIF1A_control <- HIF1A_control + plot_annotation(title = "HIF1A Coverage Plot with Peak Links - Control")

# Save the combined plots
ggsave(filename = "/home/tjq4954/Multiome_PBMC/Visuals/CoveragePlot_HIF1A_CaseLinks.png", plot = HIF1A_case, width = 9, height = 15, dpi = 600)
ggsave(filename = "/home/tjq4954/Multiome_PBMC/Visuals/CoveragePlot_HIF1A_ControlLinks.png", plot = HIF1A_control, width = 9, height = 15, dpi = 600)

```

Volcano Plot for monocytes -- to search for interesting genes worth running coverage plots
```{r}
# BiocManager::install('EnhancedVolcano')
library(EnhancedVolcano)

pbmc.rna <- readRDS("/home/tjq4954/Multiome_PBMC/Data/rna/pbmc.rna.rds")

# Remove platelets and erythrocytes
Idents(pbmc.rna) <- "predicted.celltype.l2"
pbmc.rna <- subset(pbmc.rna, idents = c("Eryth", "Platelet"), invert = TRUE)
unique(pbmc.rna$predicted.celltype.l2)

# Exclude "RNA_161" from the dataset
# pbmc.rna <- subset(pbmc.rna, subset = orig.ident != "RNA_161") # leave out the bad sample
# unique(pbmc.rna$orig.ident) # check the remaining samples
Idents(pbmc.rna) <- "predicted.celltype.l1"
pbmc_subset <- subset(pbmc.rna, idents = c("Mono")) # only include the monocyte population

# Subsetting case v.s. control
control_samples <- c("RNA_280", "RNA_302")
case_samples <- c("RNA_72", "RNA_203", "RNA_271", "RNA_294")
pbmc_subset@meta.data$dataset <- ifelse(pbmc_subset@meta.data$orig.ident %in% control_samples, "control", "case")
pbmc_subset <- PrepSCTFindMarkers(pbmc_subset)
Idents(pbmc_subset) <- "dataset"
table(pbmc_subset$dataset)

markers <- FindMarkers(
  object = pbmc_subset,
  ident.1 = "case",
  ident.2 = "control",
  min.pct = 0.01,
  assay = "SCT"
)

A <- EnhancedVolcano(markers,
                rownames(markers),
                x ="avg_log2FC", 
                y ="p_val_adj",
                title = "Volcano plot of Case (R) v. Control (L) Monocytes",
                xlim = c(-5,5))

AA <- EnhancedVolcano(markers,
                rownames(markers),
                x ="avg_log2FC", 
                y ="p_val_adj",
                title = "Volcano plot of Case (R) v. Control (L) Monocytes",
                lab=NA,
                xlim = c(-5,5))


# Define file path and name for the saved plot
output_file_A <- "/home/tjq4954/Multiome_PBMC/Visuals/VolcanoPlot_Monocytes.png"
output_file_AA <- "/home/tjq4954/Multiome_PBMC/Visuals/VolcanoPlot_Monocytes_NoLabels.png"

# Save the plot
ggsave(filename = output_file_A, plot = A, width = 10, height = 8, dpi = 600)
ggsave(filename = output_file_AA, plot = AA, width = 10, height = 8, dpi = 600)

```

Volcano Plot for CD16 & CD14 monocytes -- to search for interesting genes worth running coverage plots
```{r}
# Now, add additional volcano plots tailored for cd16 & cd14 monocytes
Idents(pbmc.rna) <- "predicted.celltype.l2"
pbmc_cd16 <- subset(pbmc.rna, idents = c("CD16 Mono"))
pbmc_cd14 <- subset(pbmc.rna, idents = c("CD14 Mono"))

# Subsetting case v.s. control
control_samples <- c("RNA_280", "RNA_302")
case_samples <- c("RNA_72", "RNA_203", "RNA_271", "RNA_294")
pbmc_cd16@meta.data$dataset <- ifelse(pbmc_cd16@meta.data$orig.ident %in% control_samples, "control", "case")
pbmc_cd14@meta.data$dataset <- ifelse(pbmc_cd14@meta.data$orig.ident %in% control_samples, "control", "case")
pbmc_cd16 <- PrepSCTFindMarkers(pbmc_cd16)
pbmc_cd14 <- PrepSCTFindMarkers(pbmc_cd14)
Idents(pbmc_cd16) <- "dataset"
Idents(pbmc_cd14) <- "dataset"
table(pbmc_cd16$dataset)
table(pbmc_cd14$dataset)

markers_cd16 <- FindMarkers(
  object = pbmc_cd16,
  ident.1 = "case",
  ident.2 = "control",
  min.pct = 0.01,
  assay = "SCT"
)
markers_cd14 <- FindMarkers(
  object = pbmc_cd14,
  ident.1 = "case",
  ident.2 = "control",
  min.pct = 0.01,
  assay = "SCT"
)

markers_cd14 %>%
  top_n(-50, avg_log2FC)

B <- EnhancedVolcano(markers_cd16,
                rownames(markers_cd16),
                x ="avg_log2FC", 
                y ="p_val_adj",
                title = "Volcano plot of Case (R) v. Control (L) CD16 Monocytes",
                xlim = c(-5,5))

BB <- EnhancedVolcano(markers_cd16,
                rownames(markers_cd16),
                x ="avg_log2FC", 
                y ="p_val_adj",
                title = "Volcano plot of Case (R) v. Control (L) CD16 Monocytes",
                lab=NA,
                xlim = c(-5,5))

C <- EnhancedVolcano(markers_cd14,
                rownames(markers_cd14),
                x ="avg_log2FC", 
                y ="p_val_adj",
                title = "Volcano plot of Case (R) v. Control (L) CD14 Monocytes",
                xlim = c(-5,5))

CC <- EnhancedVolcano(markers_cd14,
                rownames(markers_cd14),
                x ="avg_log2FC", 
                y ="p_val_adj",
                title = "Volcano plot of Case (R) v. Control (L) CD14 Monocytes",
                lab=NA,
                xlim = c(-5,5))

# Save the plot
output_file_B <- "/home/tjq4954/Multiome_PBMC/Visuals/VolcanoPlot_CD16Monocytes.png"
ggsave(filename = output_file_B, plot = B, width = 10, height = 8, dpi = 600)
output_file_BB <- "/home/tjq4954/Multiome_PBMC/Visuals/VolcanoPlot_CD16Monocytes_NoLabels.png"
ggsave(filename = output_file_BB, plot = BB, width = 10, height = 8, dpi = 600)

output_file_C <- "/home/tjq4954/Multiome_PBMC/Visuals/VolcanoPlot_CD14Monocytes.png"
ggsave(filename = output_file, plot = C, width = 10, height = 8, dpi = 600)
output_file_CC <- "/home/tjq4954/Multiome_PBMC/Visuals/VolcanoPlot_CD14Monocytes_NoLabels.png"
ggsave(filename = output_file_CC, plot = CC, width = 10, height = 8, dpi = 600)

```

Violin Plot
```{r}
# Testing
VlnPlot(pbmc_case, features = c("VCAM1", "ICAM1", "CCL5", "S100A4", "ANXA1", "CCR7", "ISG15", "CD3D"),
        group.by = "predicted.celltype.l1",
        pt.size = 0.2, ncol = 4)

# Create violin plots for selected genes in the case and control groups
genes_of_interest <- c("PSMB8-AS1", "PSMB9", "ICAM1", "VCAM1", "IL1B", "HIF1A")
DefaultAssay(pbmc_case) <- "SCT"
DefaultAssay(pbmc_control) <- "SCT"

# Generate violin plots for the case group
violin_case <- VlnPlot(
  object = pbmc_case,
  features = genes_of_interest,
  pt.size = 0,
  ncol = 3,
  group.by = "predicted.celltype.l1",
  slot = 'data')

# Generate violin plots for the control group
violin_control <- VlnPlot(
  object = pbmc_control,
  features = genes_of_interest,
  pt.size = 0,
  ncol = 3,
  group.by = "predicted.celltype.l1",
  slot = 'data')

# Print violin plots
violin_case
violin_control

# Save the plots if needed
setwd("/home/tjq4954/Multiome_PBMC/Visuals")
ggsave("violin_case_plots.png", plot = violin_case, width = 15, height = 10, dpi = 600)
ggsave("violin_control_plots.png", plot = violin_control, width = 15, height = 10, dpi = 600)

```

```{r}
sessionInfo()
```
