---
title: "10XMultiome"
output: html_document
date: "2024-01-04"
---

Loading libraries
```{r}
library(Signac)
library(Seurat)
library(SeuratDisk)
library(SeuratObject)
library(SeuratData)
library(GenomicRanges)
library(future)
library(EnsDb.Hsapiens.v86)
library(Azimuth)
library(BSgenome.Hsapiens.UCSC.hg38)

# setRepositories(ind = 1:3, addURLs = c('https://satijalab.r-universe.dev', 'https://bnprks.r-universe.dev/'))
# install.packages(c("BPCells", "presto", "glmGamPoi"))
# BiocManager::install("TFBSTools", type = "source", force = TRUE)

plan("multisession", workers = 115)
options(future.globals.maxSize = 100000 * 1024^2) # for 100 Gb RAM
```

Merging objects
```{r}
# Load the ATAC and RNA objects
pbmc.atac <- readRDS("/home/tjq4954/Multiome_PBMC/Data/atac/pbmc.atac.rds")
pbmc.rna <- readRDS("/home/tjq4954/Multiome_PBMC/Data/rna/pbmc.rna.rds")

# Check if the cell names are formatted the same
identical(colnames(pbmc.atac), colnames(pbmc.rna))

# Merging objects
pbmc.rna[['ATAC']] <- pbmc.atac[['ATAC']] # assay
pbmc.rna[['lsi']] <- pbmc.atac[['lsi']] # embedding
pbmc.rna@meta.data <- cbind(pbmc.rna@meta.data, pbmc.atac@meta.data)
pbmc <- pbmc.rna

# Specify the full path with the filename for the RDS file
save_path <- "/home/tjq4954/Multiome_PBMC/Data/pbmc.rds"
# Save the Seurat object as an RDS file
saveRDS(pbmc, file = save_path)
```

Quality control
```{r}
setwd("/home/tjq4954/Multiome_PBMC/Data")
pbmc <- readRDS("pbmc.rds")

DefaultAssay(pbmc) <- "ATAC"

# Create some important QC metrics
pbmc <- NucleosomeSignal(pbmc)
pbmc <- TSSEnrichment(pbmc)

VlnPlot(
  object = pbmc,
  features = c("nCount_RNA", "nCount_ATAC","TSS.enrichment","nucleosome_signal","percent.mt"),
  ncol = 3,
  pt.size = 0
)

# filter out low quality cells
pbmc <- subset(
  x = pbmc,
  subset = nCount_ATAC < 100000 &
    nCount_RNA < 25000 &
    nCount_ATAC > 1000 &
    nCount_RNA > 1000 &
    nucleosome_signal < 2 &
    TSS.enrichment > 1
)

# Create the contingency table
Idents(pbmc) <- "predicted.celltype.l2"
celltype_ident_table <- table(pbmc$predicted.celltype.l2, pbmc$orig.ident)
celltype_ident_df <- as.data.frame(celltype_ident_table)

# Save the table as a CSV file
output_file_path <- "/home/tjq4954/Multiome_PBMC/Visuals/celltype_ident_table.csv"
write.csv(celltype_ident_df, file = output_file_path, row.names = FALSE)

# Specify the full path with the filename for the RDS file
save_path <- "/home/tjq4954/Multiome_PBMC/Data/pbmc_subset.rds"
# Save the Seurat object as an RDS file
saveRDS(pbmc, file = save_path)
```

Peak calling
```{r}
# devtools::install_github("GreenleafLab/ArchR", ref="master", repos = BiocManager::repositories())
# library(ArchR)
# ArchR::installExtraPackages()
 
setwd("/home/tjq4954/Multiome_PBMC/Data")
pbmc <- readRDS("pbmc_subset.rds")

# # call peaks using MACS
# # findMacs2()
# peaks <- CallPeaks(pbmc)
# 
# # remove peaks on nonstandard chromosomes and in genomic blacklist regions
# peaks <- keepStandardChromosomes(peaks, pruning.mode = "coarse")
# peaks <- subsetByOverlaps(x = peaks, ranges = blacklist_hg38_unified, invert = TRUE)
# 
# # quantify counts in each peak
# macs2_counts <- FeatureMatrix(
#   fragments = Fragments(pbmc),
#   features = peaks,
#   cells = colnames(pbmc)
# )
# 
# # create a new assay using the MACS2 peak set and add it to the Seurat object
# pbmc[["peaks"]] <- CreateChromatinAssay(
#   counts = macs2_counts,
#   fragments = fragpath,
#   annotation = annotation
# )
```

DEG & DA Analysis
```{r}
# DefaultAssay(pbmc) <- "RNA" # already performed in the RNA integration script
# pbmc <- SCTransform(pbmc)
# pbmc <- RunPCA(pbmc)

# DefaultAssay(pbmc) <- "peaks"
# pbmc <- FindTopFeatures(pbmc, min.cutoff = 5)
# pbmc <- RunTFIDF(pbmc)
# pbmc <- RunSVD(pbmc)
```

Joint UMAP Visualization
```{r}
library(ggplot2)

# set a reasonable order for cell types to be displayed when plotting
# levels(pbmc) <- c("CD4 Naive", "CD4 TCM", "CD4 CTL", "CD4 TEM", "CD4 Proliferating",
#                   "CD8 Naive", "dnT",
#                  "CD8 TEM", "CD8 TCM", "CD8 Proliferating", "MAIT", "NK", "NK_CD56bright",
#                  "NK Proliferating", "gdT",
#                  "Treg", "B naive", "B intermediate", "B memory", "Plasmablast",
#                  "CD14 Mono", "CD16 Mono",
#                  "cDC1", "cDC2", "pDC", "HSPC", "Eryth", "ASDC", "ILC", "Platelet")

# build a joint neighbor graph using both assays
pbmc <- FindMultiModalNeighbors(
  object = pbmc,
  reduction.list = list("pca", "lsi"), 
  dims.list = list(1:50, 2:40),
  modality.weight.name = "RNA.weight",
  verbose = TRUE
)

# build a joint UMAP visualization
pbmc <- RunUMAP(
  object = pbmc,
  nn.name = "weighted.nn",
  assay = "RNA",
  verbose = TRUE
)

# WNN UMAP - Cell Type
plot1 <- DimPlot(pbmc, label = TRUE, repel = TRUE, reduction = "umap", group.by = "predicted.celltype.l2") + NoLegend()
ggsave(filename = "/home/tjq4954/Multiome_PBMC/Visuals/JointUMAP_CellType.png", plot = plot1, width = 10, height = 8, dpi = 600)
# WNN UMAP - Sample ID
plot2 <- DimPlot(pbmc, label = TRUE, repel = TRUE, reduction = "umap", group.by = "orig.ident") + NoLegend()
ggsave(filename = "/home/tjq4954/Multiome_PBMC/Visuals/JointUMAP_SampleID.png", plot = plot2, width = 10, height = 8, dpi = 600)
```

Linking peaks to genes of interest
```{r}
# DefaultAssay(pbmc) <- "peaks"
# 
# # first compute the GC content for each peak
# pbmc <- RegionStats(pbmc, genome = BSgenome.Hsapiens.UCSC.hg38)
# 
# # link peaks to genes
# pbmc <- LinkPeaks(
#   object = pbmc,
#   peak.assay = "peaks",
#   expression.assay = "SCT",
#   genes.use = c("LYZ", "MS4A1")
# )
```

Volcano Plot for monocytes -- to search for interesting genes worth running coverage plots
```{r}
# BiocManager::install('EnhancedVolcano')
library(EnhancedVolcano)

pbmc.rna <- readRDS("/home/tjq4954/Multiome_PBMC/Data/rna/pbmc.rna.rds")

# Exclude "RNA_161" from the dataset
# pbmc.rna <- subset(pbmc.rna, subset = orig.ident != "RNA_161") # leave out the bad sample
# unique(pbmc.rna$orig.ident) # check the remaining samples
Idents(pbmc.rna) <- "predicted.celltype.l1"
pbmc_subset <- subset(pbmc.rna, idents = c("Mono")) # only include the monocyte population

# Subsetting case v.s. control
control_samples <- c("RNA_280", "RNA_302")
case_samples <- c("RNA_72", "RNA_203", "RNA_271", "RNA_294")
pbmc_subset@meta.data$dataset <- ifelse(pbmc_subset@meta.data$orig.ident %in% control_samples, "control", "case")
pbmc_subset <- PrepSCTFindMarkers(pbmc_subset)
Idents(pbmc_subset) <- "dataset"
table(pbmc_subset$dataset)

markers <- FindMarkers(
  object = pbmc_subset,
  ident.1 = "case",
  ident.2 = "control",
  min.pct = 0.01,
  assay = "SCT"
)

A <- EnhancedVolcano(markers,
                rownames(markers),
                x ="avg_log2FC", 
                y ="p_val_adj",
                title = "Volcano plot of Case (R) v. Control (L) Monocytes",
                xlim = c(-5,5))

# Define file path and name for the saved plot
output_file <- "/home/tjq4954/Multiome_PBMC/Visuals/VolcanoPlot_Monocytes.png"

# Save the plot
ggsave(filename = output_file, plot = A, width = 10, height = 8, dpi = 600)
```

Volcano Plot for CD16 & CD14 monocytes -- to search for interesting genes worth running coverage plots
```{r}
# Now, add additional volcano plots tailored for cd16 & cd14 monocytes
Idents(pbmc.rna) <- "predicted.celltype.l2"
pbmc_cd16 <- subset(pbmc.rna, idents = c("CD16 Mono"))
pbmc_cd14 <- subset(pbmc.rna, idents = c("CD14 Mono"))

# Subsetting case v.s. control
control_samples <- c("RNA_280", "RNA_302")
case_samples <- c("RNA_72", "RNA_203", "RNA_271", "RNA_294")
pbmc_cd16@meta.data$dataset <- ifelse(pbmc_cd16@meta.data$orig.ident %in% control_samples, "control", "case")
pbmc_cd14@meta.data$dataset <- ifelse(pbmc_cd14@meta.data$orig.ident %in% control_samples, "control", "case")
pbmc_cd16 <- PrepSCTFindMarkers(pbmc_cd16)
pbmc_cd14 <- PrepSCTFindMarkers(pbmc_cd14)
Idents(pbmc_cd16) <- "dataset"
Idents(pbmc_cd14) <- "dataset"
table(pbmc_cd16$dataset)
table(pbmc_cd14$dataset)

markers_cd16 <- FindMarkers(
  object = pbmc_cd16,
  ident.1 = "case",
  ident.2 = "control",
  min.pct = 0.01,
  assay = "SCT"
)
markers_cd14 <- FindMarkers(
  object = pbmc_cd14,
  ident.1 = "case",
  ident.2 = "control",
  min.pct = 0.01,
  assay = "SCT"
)

markers_cd14 %>%
  top_n(-50, avg_log2FC)

B <- EnhancedVolcano(markers_cd16,
                rownames(markers_cd16),
                x ="avg_log2FC", 
                y ="p_val_adj",
                title = "Volcano plot of Case (R) v. Control (L) CD16 Monocytes",
                xlim = c(-5,5))

C <- EnhancedVolcano(markers_cd14,
                rownames(markers_cd14),
                x ="avg_log2FC", 
                y ="p_val_adj",
                title = "Volcano plot of Case (R) v. Control (L) CD14 Monocytes",
                xlim = c(-5,5))

# Save the plot
output_file <- "/home/tjq4954/Multiome_PBMC/Visuals/VolcanoPlot_CD16Monocytes.png"
ggsave(filename = output_file, plot = B, width = 10, height = 8, dpi = 600)
output_file <- "/home/tjq4954/Multiome_PBMC/Visuals/VolcanoPlot_CD14Monocytes.png"
ggsave(filename = output_file, plot = C, width = 10, height = 8, dpi = 600)

```

Coverage Plot
```{r}
DefaultAssay(pbmc) <- "ATAC"

# first compute the GC content for each peak
# pbmc <- RegionStats(pbmc, genome = BSgenome.Hsapiens.UCSC.hg38)

# link peaks to genes
# pbmc <- LinkPeaks(
#   object = pbmc,
#   peak.assay = "peaks",
#   expression.assay = "SCT",
#   genes.use = c("LYZ", "MS4A1")
# )

# Look up genes to plot
grep("VEG", rownames(pbmc@assays$SCT), value = TRUE)

p1 <- CoveragePlot(
  object = pbmc,
  region = "ZFAND6",
  features = "ZFAND6",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l2",
  idents = c("CD16 Mono","CD14 Mono")
)

p2 <- CoveragePlot(
  object = pbmc,
  region = "IFNGR1",
  features = "IFNGR1",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l2",
  idents = c("CD16 Mono","CD14 Mono"),
  extend.downstream = 5000
)

p3 <- CoveragePlot(
  object = pbmc,
  region = "FOS",
  features = "FOS",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l2",
  idents = c("CD16 Mono","CD14 Mono")
)

p4 <- CoveragePlot(
  object = pbmc,
  region = "HIF1A",
  features = "HIF1A",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l2",
  idents = c("CD16 Mono","CD14 Mono")
  # , extend.upstream = 5000,
  # extend.downstream = 5000
)

patchwork::wrap_plots(p1,p2, ncol = 1)
patchwork::wrap_plots(p3,p4, ncol = 1)

```

Cases vs. Controls for IL1B, CCL4, MIF, HIF1A
```{r}
# Assuming you have already identified which samples are cases and controls
control_samples <- c("RNA_280", "RNA_302")
case_samples <- c("RNA_72", "RNA_203", "RNA_271", "RNA_294")

# Subset pbmc object into case and control groups
pbmc_control <- subset(pbmc, subset = orig.ident %in% control_samples)
pbmc_case <- subset(pbmc, subset = orig.ident %in% case_samples)

# Look up genes to plot
grep("SPP", rownames(pbmc_case@assays$SCT), value = TRUE)

# Coverage plots for case group
p1_case <- CoveragePlot(
  object = pbmc_case,
  region = "IL1B",
  features = "IL1B",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l2",
  idents = c("CD16 Mono", "CD14 Mono")
)

p2_case <- CoveragePlot(
  object = pbmc_case,
  region = "CD36",
  features = "CD36",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l2",
  idents = c("CD16 Mono", "CD14 Mono")
)

p3_case <- CoveragePlot(
  object = pbmc_case,
  region = "CXCR4",
  features = "CXCR4",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l2",
  idents = c("CD16 Mono", "CD14 Mono")
)

p4_case <- CoveragePlot(
  object = pbmc_case,
  region = "HIF1A",
  features = "HIF1A",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l2",
  idents = c("CD16 Mono", "CD14 Mono")
)

# Coverage plots for control group
p1_control <- CoveragePlot(
  object = pbmc_control,
  region = "IL1B",
  features = "IL1B",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l2",
  idents = c("CD16 Mono", "CD14 Mono")
)

p2_control <- CoveragePlot(
  object = pbmc_control,
  region = "CD36",
  features = "CD36",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l2",
  idents = c("CD16 Mono", "CD14 Mono")
)

p3_control <- CoveragePlot(
  object = pbmc_control,
  region = "CXCR4",
  features = "CXCR4",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l2",
  idents = c("CD16 Mono", "CD14 Mono")
)

p4_control <- CoveragePlot(
  object = pbmc_control,
  region = "HIF1A",
  features = "HIF1A",
  expression.assay = "SCT",
  group.by = "predicted.celltype.l2",
  idents = c("CD16 Mono", "CD14 Mono")
)

# Add titles to case and control coverage plots
p1_case <- p1_case + plot_annotation(title = "IL1B - Case (Top) v.s. Control (Bottom)")
p1_control <- p1_control
p2_case <- p2_case + plot_annotation(title = "CD36 - Case (Top) v.s. Control (Bottom)")
p2_control <- p2_control
p3_case <- p3_case + plot_annotation(title = "CXCR4 - Case (Top) v.s. Control (Bottom)")
p3_control <- p3_control
p4_case <- p4_case + plot_annotation(title = "HIF1A - Case (Top) v.s. Control (Bottom)")
p4_control <- p4_control

# Combine case and control plots for each gene
library(patchwork)
combined_p1 <- p1_case + p1_control + plot_layout(ncol = 1)
combined_p2 <- p2_case + p2_control + plot_layout(ncol = 1)
combined_p3 <- p3_case + p3_control + plot_layout(ncol = 1)
combined_p4 <- p4_case + p4_control + plot_layout(ncol = 1)

# Save the combined plots
ggsave(filename = "/home/tjq4954/Multiome_PBMC/Visuals/CoveragePlot_IL1B.png", plot = combined_p1, width = 10, height = 8, dpi = 600)
ggsave(filename = "/home/tjq4954/Multiome_PBMC/Visuals/CoveragePlot_CD36.png", plot = combined_p2, width = 10, height = 8, dpi = 600)
ggsave(filename = "/home/tjq4954/Multiome_PBMC/Visuals/CoveragePlot_CXCR4.png", plot = combined_p3, width = 10, height = 8, dpi = 600)
ggsave(filename = "/home/tjq4954/Multiome_PBMC/Visuals/CoveragePlot_HIF1A.png", plot = combined_p4, width = 10, height = 8, dpi = 600)

table(pbmc$orig.ident)
```

Violin Plot
```{r}
# Create violin plots for selected genes in the case and control groups
genes_of_interest <- c("IL1B", "CD36", "CXCR4", "HIF1A")
DefaultAssay(pbmc_case) <- "SCT"
DefaultAssay(pbmc_control) <- "SCT"

# Generate violin plots for the case group
violin_case <- VlnPlot(
  object = pbmc_case,
  features = genes_of_interest,
  pt.size = 0,
  ncol = 2,
  group.by = "predicted.celltype.l2"
)

# Generate violin plots for the control group
violin_control <- VlnPlot(
  object = pbmc_control,
  features = genes_of_interest,
  pt.size = 0,
  ncol = 2,
  group.by = "predicted.celltype.l2"
)

# Print violin plots
violin_case
violin_control

# Save the plots if needed
setwd("/home/tjq4954/Multiome_PBMC/Visuals")
ggsave("violin_case_plots.png", plot = violin_case, width = 12, height = 6, dpi = 600)
ggsave("violin_control_plots.png", plot = violin_control, width = 12, height = 6, dpi = 600)
```

```{r}
sessionInfo()
```
